Open-Set Domain Adaptation through Self-Supervision

### 2022-02-11 paper/main.tex
3.3 Rotation Recognition reviewed

Summarizing, in OSDA tasks, we have both a domain gap ($p_s \neq p_t^{\mathcal{C}_s}$)
and a category gap ($\mathcal{C}_s \neq \mathcal{C}_t$). Moreover, the goal is to assign the target samples either to a category ${i \in \mathcal{C}_s}$,
or to reject them as {\it unknown}.
A metric to measure the complexity of the OSDA problem is the {\it openness} betweeen the source and the target domain \cite{bendale2015open}, 
defined as $\displaystyle \mathbb{O} = 1-\frac{\mathcal{C}_s}{\mathcal{C}_t}$.
When $\mathbb{O} > 0$, we are dealing with an OSDA problem, otherwise we are in a in a CSDA setting.

@@

We denote with $rot({\bf x}, k)$ the rotation of the sample image {\bf x} by $k\times 90$ degrees clockwise.
This is the self-supervised part of our proposed model as rotation indices $k \in [0, 3]$ can be randomly generated and then predicted.
Relative rotation,
meaning that both original and rotated features are supplied to the model,
is preferred over just using rotated features as some objects might not have an absolute correct orientation.
Alternatively, 
instead of having the model predict the rotation of a sample for any class using the same classifier, 
we also try using a different head for each one of the known classes $\mathcal{C}_s$.

>>>>

Summarizing, in OSDA tasks, we have both a domain gap ($p_s \neq p_t^{\mathcal{C}_s}$)
and a category gap ($\mathcal{C}_s \neq \mathcal{C}_t$). Moreover, the goal is to assign the target samples either to a category ${i \in \mathcal{C}_s}$,
or to reject them as {\it unknown}.
A metric to measure the complexity of the OSDA problem is the {\it openness} betweeen the source and the target domain \cite{bendale2015open}, 
defined as $\displaystyle \mathbb{O} = 1-\frac{\mathcal{C}_s}{\mathcal{C}_t}$.
When $\mathbb{O} > 0$, we are dealing with an OSDA problem, otherwise we are in a CSDA setting.

@@

We denote with $rot({\bf x}, k)$ the rotating function of a sample image {\bf x} by $k\times 90$ degrees clockwise.
The self-supervised pre-text task consists in generating a random rotation index $k \in [0, 3]$ that then becomes
the label for the rotated version of the image ${\bf \tilde{x}} = rot({\bf x}, k)$.
Then, the task becomes a standard classification of the correct rotation index ($\mathcal{C}_r = \{0, 1, 2, 3\}$).

{\it Relative orientation:}
more precisely, we exploit a relative rotation task,
implying that both the features of the original and rotated image are supplied to the rotation classifier.
This is preferred over the absolute rotation task as some objects might not have an absolute coherent orientation
inside the dataset
(e.g. a pen may be present in different rotated versions inside the dataset).

{\it Multi-head rotation classifier:}
alternatively, 
instead of having a single rotation head predicting the rotation of a sample regardless of its semantic class, 
we also try using a different head for each known class $\in \mathcal{C}_s$, each one responsible of predicting the rotation
of the images belonging to that semantic class.
This variation can mitigate the problem of trying to predict the rotation of a larger number
of semantic classes. Infact, as the number of semantic classes grows,
the problem of predicting the relative orientation becomes more difficult.
The application of the rotation recognition pre-text task allows to effectively favor and force the model to learn
domain-independent patterns, which are crucial to perform the novelty detection in a cross-domain fashion and, moreover,
to successfully perform the domain alignment.
To provide an explanation of why this applies, we can think that a rotation classifier needs to focus on discriminative patterns to
successfully perform the rotation predictions, such as shapes, edges, and high-level object relative position like the position of the eyes
w.r.t nose.
The method and its effectiveness is further illustrated in \cite{OldROS}. 
We will further discuss possible improvements and variants in {\bf PLEASE COMPLETE}.
